Response to Reviewers

### 2025-07-08 paper/kobakian-cook-anzjs.tex
typo in equation fixed

The Australian Cancer Atlas \citep{atlas} is an online interactive web
tool created to explore the burden of cancer on Australian communities.
There are many cancer types to be explored individually or aggregated.
The Australian Cancer Atlas allows users to explore the patterns in the
distributions of cancer statistics over the geographic space of
Australia. It uses a choropleth map display and diverging color scheme
to draw attention to relationships between neighboring areas. The
hexagon tile map may be a useful alternative display to enhance the
atlas.

The experiment was conducted using the lineup protocol, a visual
inference procedure \citep{GIIV}, to objectively test the effectiveness
of the two displays.

@@

Spatial visualisations communicate the distribution of statistics over
geographic landscapes. The choropleth map \citep[\citet{BCM}]{EI} is a
traditional display. It is used to present statistics that have been
aggregated on geographic units. Creating a choropleth map involves
drawing polygons representing the administrative boundaries, and filling
with colour mapped to the value of the statistic. The choropleth map
places the statistic in the context of the spatial domain, so that the
reader can see whether there are spatial trends, clusters or anomalies.
This is important for digesting disease patterns. If there is a trend it
may imply that the disease is spreading from one location to another. If
there is a cluster, or an anomaly, there may be a localized outbreak of
the disease. Aggregating the statistic on administrative units, provides
a level of privacy to individuals, while allowing the impact of the
disease on the community to be analyzed.

@@

A cartogram is a general solution for adequately displaying a
population-based statistic. It transforms the geographic map base to
reflect the population in the geographic region, while preserving some
aspects of the geographic location. There are several cartogram
algorithms \citep[\citet{CBATCC}]{ACTUC}; each involves shifting the
boundaries of geographic units, using the value of the statistic to
increase or decrease the area taken by the geographic unit on the map.
The changes to the boundaries result in cartograms that accurately
communicate population by map area for each of the geographic units but
can result in losing the familiar geographic information. For Australia,
the transformations warp the country so that it is no longer
recognizable.

Alternative algorithms make various trade offs between familiar shapes
and representation of geographic units. The non-contiguous cartogram
method \citep{NAC} keeps the shapes of geographic units intact, and
changes the size of the shape. This method disconnects areas creating
empty space on the display losing the continuity of the spatial display
of the statistic. The Dorling cartogram \citep{ACTUC} represents each
unit as a circle, sized according to the value of the statistic. The
neighbour relationships are mostly maintained by how the circles touch.
A similar approach was pioneered by \citet{RSCW}, using rectangles that
tile to align borders of neighbours \citep{CDWCS}. There have been
thorough reviews of the array of methods, as suitable for cancer atlas
displays \citep[\citet{BCM}]{review}.

The hexagon tile map algorithm, automatically matches spatial regions to
their nearest hexagon tile, from a grid of tiles. It has the effect of
spreading out the inner city areas while maintaining the spatial
locations or regions in remote areas. The algorithm is available in the
R package, sugarbag \citep{sugarbag}. Figure~\ref{fig-thyroid} b shows
the hexagon tile map, where the map is coloured from substantially below
average (blue) to substantially above average (red) rates. The inner
city areas have expanded, making it possible to see the cancer incidence
in the small, densely populated areas. Remote regions are represented by
isolated hexagons, which is not ideal, but maintains the spatial
location of these data values. It is of interest to know how well the
spatial distribution is perceived for this display, in comparison to the
choropleth.

@@

In order to assess the effectiveness of the hexagon tile map, the lineup
protocol \citep[\citet{GIIV}]{BCHLLSW09} from visual inference
procedures is employed. The approach mirrors classical statistical
inference. The procedures for doing a power comparison of competing plot
designed, outlined in \citet{GTPCCD}, are followed.

@@

Australia is used for the study, with Statistical Area 3 (SA3)
\citep{abs2016} as the geographic units. The results should apply
broadly to any other geographic area of interest.

@@

\begin{itemize}
\tightlist
\item
  Map type: \emph{Choropleth, Hexagon tile}
\item
  Trend: \emph{South-East to North-West; Locations in three population
  centres; Locations in multiple population centres, }
\end{itemize}

@@

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{kobakian-cook-anzjs_files/figure-pdf/fig-exp-design-1.pdf}

@@

\caption{\label{fig-exp-design}The experimental design used in the
study. Participants are allocated to with group A or B, to evaluate
either the choropleth or hexagon tile map lineup of each simulated data
set. Place in the lineup is recorded as `Location'.}

@@

The first step in the data cleaning process involved checking that
survey responses collected for each participants were only included once
in the data set. The data cleaning process also involved filtering out
participants' who did not provide at least three unique choices when
considering each of the twelve lineups. These participants achieved a
detection rate of 0. If participants had made various plot choices for
the 12 displays they saw they were still included in the dataset.

\[
\text{logit}(p_{ijk}) = \mu_i + \tau_j + \delta_k + (\tau \delta)_{jk} ~~~~ j=1,2; ~~ k=1,2,3
\]

where \(y_{ijk} = 0, 1\) represents whether subject \(i\) detected the
data plot (1) or did not (0), \(\mu_\) is the subject-specific random
intercept, \(\tau_j\) is the map type effect, \(\delta_k\) is the trend
model effect. The interaction between map type and trend model allows
for any map type effect to differ between trend models. As each
participant provides results from 12 lineups, this model can account for
each individual participants' abilities with the subject-specific random
intercept.

@@

\begin{itemize}
\tightlist
\item
  Plot choice: the number of the plot that they deemed to be most
  different from the others.
\item
  Reason: one of ``Clusters of colour'', ``Colour trend across the
  areas'', ``Big differences between neighbouring areas'', ``All areas
  have similar colours'' or ``None of these reasons''.
\item
  Certainty: how certain that their choice is different from the others,
  on a scale of 1-5.
\end{itemize}

@@

@@

Figure~\ref{fig-detect-compare} displays the average detection rates for
the two types of plot separately for each trend model. Each trend model
was tested using four repetitions, evaluations on the same data set were
seen as either choropleths or hexagon tile maps by each group as
specified in Figure~\ref{fig-exp-design}; the detection rates for each
display are connected by a line segment. The Three Cities and All Cities
trend models shown in the hexagon tile map allowed viewers to detect the
data plot substantially more often than the choropleth counterparts. One
replicate for the All Cities group had similar detection rates for both
map types, the rate of detection using the choropleth map was much
higher than other replicates. Surprisingly, participants could also
detect the gradual spatial trend in the NW-SE group from the hexagon
tile map. We expected that the choropleth map would be superior for the
type of spatial pattern, but the data suggests the hexagon tile map
performs slightly better, or equally as well.

@@

\caption{\label{tbl-desc-stats}}

@@

\caption{\label{tab:tbl-desc-stats}The mean and standard deviation of the rate of detection for each trend model, calculated for the choropleth and hexagon tile map displays.}
\centering
\begin{tabular}[t]{lccc}
\toprule
Type & NW-SE & Three Cities & All Cities\\
\midrule
Choro. & 0.52 & 0.04 & 0.23\\
 & (0.50) & (0.19) & (0.42)\\
\addlinespace
Hex. & 0.62 & 0.40 & 0.58\\
 & (0.49) & (0.49) & (0.49)\\
\bottomrule
\end{tabular}

@@

When a choropleth map display is used, the predicted detection rate for
the Three Cities trend, 0.03; this is extremely low, especially compared
to the NW-SE trend of 0.52. When the All Cities trend is presented in a
choropleth display the predicted probability of detection is 0.22. The
hexagon tile map has a substantially high detection rate for the display
of a Three Cities trend 0.39 and All Cities trend 0.59.

@@

\caption{\label{tbl-detect-glmer1}}

@@

\caption{\label{tab:tbl-detect-glmer1}The model output for the generalised linear mixed effect model for detection rate. This model considers the type of display, the trend model hidden in the data plot, and accounts for contributor performance.}
\centering
\begin{tabular}[t]{rrlrr}
\toprule
Term & Est. & Sig. & Std. Error & P val\\
\midrule
Intercept & -1.27 & $^{***}$ & 0.19 & 0.00\\
Hex. & 1.63 & $^{***}$ & 0.24 & 0.00\\
\addlinespace
Three Cities & -2.07 & $^{***}$ & 0.43 & 0.00\\
All Cities & 1.34 & $^{***}$ & 0.24 & 0.00\\
\addlinespace
Hex:Three Cities & 1.28 & $^{**}$ & 0.48 & 0.01\\
Hex:All Cities & -1.16 & $^{***}$ & 0.33 & 0.00\\
\bottomrule
\end{tabular}

@@

Figure~\ref{fig-beeswarm} shows horizontally jittered dot plots to
contrast the time taken by participants to evaluate each lineup when
viewing each type of display. The time are also separated by trend model
and whether the data plot was detected or not detected. The time taken
to complete an evaluation ranged from milliseconds to 60 seconds. The
average time taken for type of display is shown as a large colored dot
on each plot. when considering the heights of the green and orange dots,
there is little difference in the average time taken to read a
choropleth or hexagon tile map. Comparing the same colored dot across
each trend model row, there is a slight increase in the time taken to
correctly detected the data plot in the hexagon tile map lineup, but
little difference in evaluation time for the choropleth display.
However, there were substantially less correct detections for choropleth
lineups for the Three cities and All Cities trends.

@@

\caption{\label{fig-beeswarm}The distribution of the time taken
(seconds) to submit a response for each combination of trend, whether
the data plot was detected, and type of display, shown using
horizontally jittered dotplots. The colored point indicates average time
taken for each map type. Although some participants take just a few
seconds per evaluation, and some take as much as mcuh as 60 seconds, but
there is very little difference in time taken between map types.}

@@

Participants provided their level of certainty regarding their choice
using a five point scale. Unlike the accuracy and speed of responses
that were derived during the data processing phase, this was a
subjective assessment by the participant prompted by the question: `How
certain are you about your choice?'. Figure~\ref{fig-certainty} shows
the amount of times participants provided each level of certainty. This
was separated for each combination of trend models and display type, and
colored depending on whether a participant correctly detected the data
plot in the lineup. Participants often chose 4 or 5 when viewing the
population related trends in the choropelth display, even though they
were often incorrect when viewing an All Cities trend and overwhelmingly
incorrect for the Three Cities trend. This shows overconfidence in their
detection ability when using a choropleth map display. Participants were
less likely to be certain when their choice was incorrect and they were
viewing a hexagon tile map. For each trend model, participants were more
likely to doubt their choice and choose 1 or 2 in the hexagon tile map
displays, even though many had made the correct choice.

@@

Participants were asked why they had made their plot choice and were
able to select from a set of suggested reasons. ``Color trend across the
areas'' was the most common selection for NW-SE trend displays
(Table~\ref{tbl-reason}).

The reasons chosen by participants from the list provided to them varied
more when viewing choropleth displays than the hexagon tile map. The
hexagon tile map displays resulted in ``Clusters of color'' as the most
common choice made by participants.

@@

\caption{\label{tbl-reason}}

@@

\caption{\label{tab:tbl-reason}The amount of participants that selected each reason for their choice of plot when looking at each trend model shown in Choropleth and Hexagon Tile maps. The facets show whether or not the choice was correct.}
\centering
\begin{tabular}[t]{llll}
\toprule
Trend & Detected & Choro. & Hex.\\
\midrule
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash All Cities} & Yes & clusters, consistent & clusters\\
\cmidrule{1-4}
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Three Cities} & Yes & consistent & clusters\\
\cmidrule{1-4}
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash NW-SE} & Yes & trend & clusters\\
\bottomrule
\end{tabular}

@@

The color scaling applied in Three cities and All cities displays
resulted in the rural areas of the real data plot appearing more blue or
yellow than the other plots in the lineups. Due to the consistent
coloring of rural areas in a choropleth display, the choice ``All areas
have similar colors'' was most common reason for a participants choice.
The All Cities displays colored the inner-city areas of all capital
cities more red, this was observable to participants and explains the
equal choice of the city clusters or rural color consistency. Choosing
``Clusters of colour'' was expected when participants viewed the Hexagon
tile map display of the All Cities and Three Cities distributions. It
was unexpected that it was also the most common reason for the NW-SE
hexagon tile map displays. Due to the spatial covariance introduced in
the smoothing, groups of similarly colored hexagons were present in all
of the hexagon tile map displays. All Cities and Three Cities
distributions of real data trends had distinctly different patterns or
red inner-city areas, while some of the plots in each lineup may have
shared similar features.

@@

@@

>>>>

The Australian Cancer Atlas \citep{atlas} is an online interactive web
tool created to explore the burden of cancer on Australian communities.
There are many cancer types to be explored individually or aggregated.
The Australian Cancer Atlas allows users to explore the patterns in the
distributions of cancer statistics over the geographic space of
Australia. It uses a choropleth map display and diverging color scheme
to draw attention to relationships between neighbouring areas. The
hexagon tile map may be a useful alternative display to enhance the
atlas.

The experiment was conducted using the lineup protocol, a visual
inference procedure \citep{BCHLLSW09, GIIV}, that can be used to
objectively test the effectiveness of the two displays \citep{GTPCCD}. A
lineup embeds the data plot among a field of null data plots, and an
independent observer is asked to select the most different plot. It was
shown by \citet{VVSIALM} to be an effective way to conduct a hypothesis
test where a plot is treated as a test statistic. If the data plot is
selected it is analogous to a rejection of the null hypothesis (no
structure), and the observed data plot is unlikely to arise from the
null scenario.

@@

Spatial visualisations communicate the distribution of statistics over
geographic landscapes. The choropleth map \citep{EI, BCM} is a
traditional display. It is used to present statistics that have been
aggregated on geographic units. Creating a choropleth map involves
drawing polygons representing the administrative boundaries, and filling
with colour mapped to the value of the statistic. The choropleth map
places the statistic in the context of the spatial domain, so that the
reader can see whether there are spatial trends, clusters or anomalies.
This is important for digesting disease patterns. If there is a linear
trend it may imply a relationship between disease and geographic
location. If there is a cluster, or an anomaly, there may be a localized
outbreak of the disease.

@@

A cartogram is a general solution for adequately displaying a
population-based statistic. It transforms the geographic map base to
reflect the population in the geographic region, while preserving some
aspects of the geographic location. There are several cartogram
algorithms \citep{ACTUC, CBATCC}; each involves shifting the boundaries
of geographic units, using the value of the statistic to increase or
decrease the area taken by the geographic unit on the map. The changes
to the boundaries result in cartograms that accurately communicate
population by map area for each of the geographic units but can result
in losing the familiar geographic information. For Australia, the
transformations warp the country so that it is no longer recognizable.

Alternative algorithms make various trade offs between familiar shapes
and representation of geographic units. The non-contiguous cartogram
method \citep{NAC} keeps the shapes of geographic units intact, and
changes the size of the shape. This method disconnects areas creating
empty space on the display losing the continuity of the spatial display
of the statistic. The Dorling cartogram \citep{ACTUC} represents each
unit as a circle, sized according to the value of the statistic. The
neighbour relationships are mostly maintained by how the circles touch.
A similar approach was pioneered by \citet{RSCW}, using rectangles that
tile to align borders of neighbours \citep{CDWCS}. There have been
thorough reviews of the array of methods, as suitable for cancer atlas
displays \citep{review, BCM}.

The hexagon tile map algorithm, automatically matches spatial regions to
their nearest hexagon tile, from a grid of tiles. It has the effect of
spreading out the inner city areas while maintaining the spatial
locations or regions in remote areas. The algorithm is available in the
R package, sugarbag \citep{sugarbag}. Figure~\ref{fig-thyroid} b shows
the hexagon tile map, where the map is coloured from substantially below
average (blue) to substantially above average (red) rates. The inner
city areas have expanded, making it possible to see the cancer incidence
in the small, densely populated areas. Remote regions are represented by
isolated hexagons, which is not ideal, but maintains the spatial
location of these data values. It is of interest to know how well the
spatial distribution patterns are seen from this display, in comparison
to how they are seen from the choropleth.

@@

In order to assess the effectiveness of the hexagon tile map, the lineup
protocol \citep{BCHLLSW09, GIIV} from visual inference procedures is
employed. The approach mirrors classical statistical inference. The
procedures for doing a power comparison of competing plot designed,
outlined in \citet{GTPCCD}, are followed.

@@

Australia is used for the study, with Statistical Area 3 (SA3)
\citep{abs2016} as the geographic units. The results should apply
broadly to any other geographic areas of interest, if there are large
differences in area and population size.

@@

\begin{itemize}
\tightlist
\item
  Map type: \emph{choropleth, hexagon tile}
\item
  Trend: \emph{South-East to North-West; Locations in three population
  centres; Locations in multiple population centres, }
\end{itemize}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{kobakian-cook-anzjs_files/figure-pdf/fig-exp-design-1.pdf}

@@

\caption{\label{fig-exp-design}The experimental design used in the
study. Participants are allocated to with group A or B, to evaluate
either the choropleth or hexagon tile map lineup of each simulated data
set. The text `loc' refers to the location of the data plot in the
lineup.}

@@

@@

\begin{itemize}
\tightlist
\item
  \textbf{Plot choice}: the number of the plot that they deemed to be
  most different from the others.
\item
  \textbf{Reason}: one of ``Clusters of colour'', ``Colour trend across
  the areas'', ``Big differences between neighbouring areas'', ``All
  areas have similar colours'' or ``None of these reasons''.
\item
  \textbf{Certainty}: how certain that their choice is different from
  the others, on a scale of 1-5.
\end{itemize}

@@

Data is checked to ensure that survey responses collected for each
participants were only included once. Technically it is possible to
submit results more than once if the submit button is clicked multiple
times in short sequence. Participants who did not finish the evaluation
of all lineups or clicked through without providing their evaluation are
filtered.

\[
\text{logit}(p_{ijk}) = \mu_i + \tau_j + \delta_k + (\tau \delta)_{jk}
\]

where \(y_{ijk} = 0, 1\) represents whether subject \(i\) detected the
data plot (1) or did not (0), \(\mu_i, ~i=1, \dots n\) is the
subject-specific random intercept, \(n\) is the number of subjects,
\(\tau_j, ~j=1,2\) is the map type effect, \(\delta_k, ~k=1,2,3\) is the
trend model effect. The interaction between map type and trend model
allows for any map type effect to differ between trend models. As each
participant provides results from 12 lineups, this model can account for
each individual participants' abilities with the subject-specific random
intercept.

@@

Figure~\ref{fig-detect-compare} displays the average detection rates for
the two types of plot separately for each trend model. Each trend model
was tested using four repetitions, evaluations on the same data set were
seen as either choropleths or hexagon tile maps by each group as
specified in Figure~\ref{fig-exp-design}; the detection rates for each
display are connected by a line segment. The Three Cities and All Cities
trend models shown in the hexagon tile map allowed viewers to detect the
data plot substantially more often than the choropleth counterparts. One
replicate for the All Cities group had similar detection rates for both
map types - the rate of detection using the choropleth map was much
higher than other replicates. Surprisingly, participants could also
detect the gradual spatial trend in the NW-SE group from the hexagon
tile map. We expected that the choropleth map would be superior for the
type of spatial pattern, but the data suggests the hexagon tile map
performs slightly better, or equally as well.

@@

\caption{\label{tbl-desc-stats}The mean and standard deviation of the
rate of detection for each trend model, calculated for the choropleth
and hexagon tile map displays.}

@@

\begin{tabular}{lccc}
\toprule
Type & NW-SE & Three Cities & All Cities\\
\midrule
Choro. & 0.52 & 0.04 & 0.23\\
 & (0.50) & (0.19) & (0.42)\\
\addlinespace
Hex. & 0.62 & 0.40 & 0.58\\
 & (0.49) & (0.49) & (0.49)\\
\bottomrule
\end{tabular}

@@

When a choropleth map display is used, the predicted detection rate for
the Three Cities trend, 0.03 is extremely low, especially compared to
the NW-SE trend of 0.52. When the All Cities trend is presented in a
choropleth display the predicted probability of detection is 0.22. The
hexagon tile map has a substantially high detection rate for the display
of a Three Cities trend 0.39 and All Cities trend 0.59.

@@

\caption{\label{tbl-detect-glmer1}Parameter estimates of the fitted
model fit for detection rate. All terms are statistically significant
(\(^{**}=0.01\), \(^{***}=0.001\)).}

@@

\begin{tabular}{rrlrr}
\toprule
Term & Est. & Std. Err. & P-val. & Sig.\\
\midrule
Intercept & -1.27 & 0.19 & 0.00 & $^{***}$\\
Hex. & 1.63 & 0.24 & 0.00 & $^{***}$\\
\addlinespace
Three Cities & -2.07 & 0.43 & 0.00 & $^{***}$\\
All Cities & 1.34 & 0.24 & 0.00 & $^{***}$\\
\addlinespace
Hex:Three Cities & 1.28 & 0.48 & 0.01 & $^{**}$\\
Hex:All Cities & -1.16 & 0.33 & 0.00 & $^{***}$\\
\bottomrule
\end{tabular}

@@

Figure~\ref{fig-beeswarm} shows horizontally jittered dot plots to
contrast the time taken by participants to evaluate each lineup when
viewing each type of display. The time are also separated by trend model
and whether the data plot was detected or not detected. The time taken
to complete an evaluation ranged from milliseconds to 60 seconds. The
average time taken for type of display is shown as a large coloured dot
on each plot. When considering the heights of the green and orange dots,
there is little difference in the average time taken to read a
choropleth or hexagon tile map. Comparing the same coloured dot across
each trend model row, there is a slight increase in the time taken to
correctly detected the data plot in the hexagon tile map lineup, but
little difference in evaluation time for the choropleth display.
However, there were substantially less correct detections for choropleth
lineups for the Three cities and All Cities trends.

@@

\caption{\label{fig-beeswarm}The distribution of the time taken
(seconds) to submit a response for each combination of trend, whether
the data plot was detected, and type of display, shown using
horizontally jittered dotplots. The coloured point indicates average
time taken for each map type. Although some participants take just a few
seconds per evaluation, and some take as much as 60 seconds, but there
is very little difference in time taken between map types.}

@@

Participants provided their level of certainty regarding their choice
using a five point scale. Unlike the accuracy and speed of responses
that were derived during the data processing phase, this was a
subjective assessment by the participant prompted by the question: ``How
certain are you about your choice?''. Figure~\ref{fig-certainty} shows a
mosaic plot summarising how participants reported their certainty with
their decision. The plots are made separately for each combination of
trend models and whether the data plot was detected or not. Colour
indicates the type of map display. The proportion of subjects evaluating
the choropleth version increases as certainty increases, which suggests
overconfidence in their detection ability when using a choropleth map
display. Participants were less likely to be certain when their choice
was incorrect and they were viewing a hexagon tile map.

@@

Participants were asked why they had made their plot choice and were
able to select from a set of suggested reasons. ``Colour trend across
the areas'' was the most common selection for NW-SE trend displays
(Table~\ref{tbl-reason}).

The reasons chosen by participants from the list provided to them varied
more when viewing choropleth displays than the hexagon tile map. The
hexagon tile map displays resulted in ``Clusters of colour'' as the most
common choice made by participants.

@@

\caption{\label{tbl-reason}The main reason that participants wrote for
their choice of plot when looking at each trend model shown in
choropleth and hexagon tile maps.}

@@

\begin{tabular}{llll}
\toprule
Trend & Detected & Choro. & Hex.\\
\midrule
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash All Cities} & Yes & clusters, consistent & clusters\\
\cmidrule{1-4}
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Three Cities} & Yes & consistent & clusters\\
\cmidrule{1-4}
 & No & trend & clusters\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash NW-SE} & Yes & trend & clusters\\
\bottomrule
\end{tabular}

@@

The colour scaling applied in Three cities and All cities displays
resulted in the rural areas of the real data plot appearing more blue or
yellow than the other plots in the lineups. Due to the consistent
colouring of rural areas in a choropleth display, the choice ``All areas
have similar colours'' was most common reason for a participants choice.
The All Cities displays coloured the inner-city areas of all capital
cities more red, this was observable to participants and explains the
equal choice of the city clusters or rural colour consistency. Choosing
``Clusters of colour'' was expected when participants viewed the Hexagon
tile map display of the All Cities and Three Cities distributions. It
was unexpected that it was also the most common reason for the NW-SE
hexagon tile map displays. Due to the spatial covariance introduced in
the smoothing, groups of similarly coloured hexagons were present in all
of the hexagon tile map displays. All Cities and Three Cities
distributions of real data trends had distinctly different patterns or
red inner-city areas, while some of the plots in each lineup may have
shared similar features.

@@

\section*{Supplementary materials and
reproducibility}\label{supplementary-materials-and-reproducibility}
\addcontentsline{toc}{section}{Supplementary materials and
reproducibility}

@@

\section*{Ethics Declaration}\label{ethics-declaration}
\addcontentsline{toc}{section}{Ethics Declaration}

### 2025-07-08 paper/review1/ReponseToReviewers2025.qmd
typo in equation fixed

@@

*XXX re-make figure*

@@

Line 82. The word “perceived” here made me think about who is doing the perceiving and whether a discussion somewhere about the different audiences of spatial data displays is warranted / would be helpful. For example, who are the users of the Australian Cancer Atlas?

@@

Lines 128-129. I am not sure this claim is legitimate: “The results should apply broadly to any other geographic area of interest.” What about a choropleth map in which the geographic areas are more equally sized? (Not all choropleth maps are as extreme as the SA3 Australian one!)

@@

Section 3.7, lines 291-293. Please provide more information about the nature and structure of these three questions: plot choice, reason, and difficulty. You do not provide information until Section 4, which feels too late. Additionally, should the “certainty” question, referenced in Section 4.4, also be included here? I also wonder if Section 3.7 should be renamed, given that it is about more than just demographic data collection.

@@

@@

@@

I appreciate the opportunity to review this interesting paper that describes an experimental study comparing the effectiveness of hexagon tile maps against choropleth maps using “the lineup protocol.” While choropleth maps are a widely used visualization technique, newly proposed methods of displaying spatial data, such as hexagon tile maps, are not widely known to recipients of such data displays. In this context, the paper addresses the important question of the usability of this new plot type. With the exception of Kawakami et al., 2024, which analyzes a different type of hexagon maps, there are to my knowledge no findings on the effectiveness of hexagon tile maps so far.

@@

@@

1\) The section “Visual Inference” (Page 4) covers the background of visual inference and the lineup protocol, focusing on its use as a tool for statistical inference. I would like to see the reasoning for using it as a method to compare effectiveness made more clear. From the later text, I understand the argumentation as follows: Because the data is simulated in a way that one location shows the same significant result in two different visualizations (hexagon and choropleth), the detection rate indicates how easily participants can make the detection with respect to the map type.

@@

Second: The interpretation of the interaction effects on Page 13 Line 329 is a as follows:  “Allowing for the interaction effect, the difference in detection rate decreases for population-related displays for a choropleth map lineup but increases for a hexagon tile map display.“

@@

Additionally, I can’t replicate the detection probabilities mentioned in the text, e.g., Page 13, Lines 333f.: “For the NW-SE distribution, the predicted detection rate for the hexagon tile map display increases the predicted probability of detection

to 0.63 from 0.52 for choropleths, this is almost exactly the difference seen in the table of means and is significant only at the 0.05 level.”

Are these fixed effects estimates? I can’t replicate this and the following results from Table 2, but maybe I’m missing something?

@@

Page 1 line 16: “None of the existing approaches for creating cartograms or hexagon tiling perform well for the Australian landscape”

@@

Page 3 62 “For Australia, the transformations warp the country so that it is no longer recognizable” -\> maybe cite Kobakian, Cook & Roberts 2020 to make clear why this statement is made?

Page 3 Line 78: what’s “substantially above”?

@@

Page 7 Line 164: “smoothed several times”; specify

@@

@@

Page 9 Line 201: “The first step in the data cleaning process involved checking that survey responses collected for each participant were only included once in the data set.” Why is this necessary? Without context this sentence is confusing.

@@

Page 11: Line 269 what’s the meaning of the levels of Figure Eight participants?

@@

Page 14: The section “Reason” is extremely short and might be merged with section 4.3 (Speed) on page 14.

@@

>>>>

*Fixed and checked.*

@@

*This figure has been re-designed. Location is explained in the caption, also.*

@@

*Rephrased*

Line 82. The word "perceived" here made me think about who is doing the perceiving and whether a discussion somewhere about the different audiences of spatial data displays is warranted / would be helpful. For example, who are the users of the Australian Cancer Atlas?

*This is not especially about the Atlas, albeit motivated and grounded by the Atlas. We don't think it is relevant to discuss users of the Atlas. From our perspective, though, it would be ideal if the Atlas adopted this different display, and perhaps publication of this study might help to achieve this. We have modified the text in this paragraph to put the focus on the hexagon tile map vs the choropleth.*

@@

Lines 128-129. I am not sure this claim is legitimate: "The results should apply broadly to any other geographic area of interest." What about a choropleth map in which the geographic areas are more equally sized? (Not all choropleth maps are as extreme as the SA3 Australian one!)

*Language has been modified.*

@@

*Sections have been re-organised.*

Section 3.7, lines 291-293. Please provide more information about the nature and structure of these three questions: plot choice, reason, and difficulty. You do not provide information until Section 4, which feels too late. Additionally, should the "certainty" question, referenced in Section 4.4, also be included here? I also wonder if Section 3.7 should be renamed, given that it is about more than just demographic data collection.

*Done. The section was re-named to "Data collection". "Difficulty" was replaced by "Certainty", making terminology consistent.*

@@

*Yes, but it is relatively weak. It's something found from other studies that people were often more confident with sub-optimal plots!*

@@

*Thanks and we appreciate the careful reading.*

@@

I appreciate the opportunity to review this interesting paper that describes an experimental study comparing the effectiveness of hexagon tile maps against choropleth maps using "the lineup protocol." While choropleth maps are a widely used visualization technique, newly proposed methods of displaying spatial data, such as hexagon tile maps, are not widely known to recipients of such data displays. In this context, the paper addresses the important question of the usability of this new plot type. With the exception of Kawakami et al., 2024, which analyzes a different type of hexagon maps, there are to my knowledge no findings on the effectiveness of hexagon tile maps so far.

@@

*Thank you.*

@@

1\) The section "Visual Inference" (Page 4) covers the background of visual inference and the lineup protocol, focusing on its use as a tool for statistical inference. I would like to see the reasoning for using it as a method to compare effectiveness made more clear. From the later text, I understand the argumentation as follows: Because the data is simulated in a way that one location shows the same significant result in two different visualizations (hexagon and choropleth), the detection rate indicates how easily participants can make the detection with respect to the map type.

@@

Second: The interpretation of the interaction effects on Page 13 Line 329 is a as follows: "Allowing for the interaction effect, the difference in detection rate decreases for population-related displays for a choropleth map lineup but increases for a hexagon tile map display."

@@

Additionally, I can’t replicate the detection probabilities mentioned in the text, e.g., Page 13, Lines 333f.: "For the NW-SE distribution, the predicted detection rate for the hexagon tile map display increases the predicted probability of detection

to 0.63 from 0.52 for choropleths, this is almost exactly the difference seen in the table of means and is significant only at the 0.05 level."

Are these fixed effects estimates? I can't replicate this and the following results from Table 2, but maybe I’m missing something?

@@

Page 1 line 16: "None of the existing approaches for creating cartograms or hexagon tiling perform well for the Australian landscape"

@@

Page 3 62 "For Australia, the transformations warp the country so that it is no longer recognizable" -\> maybe cite Kobakian, Cook & Roberts 2020 to make clear why this statement is made?

Page 3 Line 78: what's "substantially above"?

@@

Page 7 Line 164: "smoothed several times"; specify

*Re-worded.*

@@

*Re-worded.*

@@

Page 9 Line 201: "The first step in the data cleaning process involved checking that survey responses collected for each participant were only included once in the data set." Why is this necessary? Without context this sentence is confusing.

@@

Page 11: Line 269 what's the meaning of the levels of Figure Eight participants?

@@

Page 14: The section "Reason" is extremely short and might be merged with section 4.3 (Speed) on page 14.

@@

*Fixed.*
